---
title: "R4DS_2E_databases"
author: "Martin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DBI)
library(dbplyr)
library(tidyverse)
library(duckdb)
```

## Databases - SQL

```{r Database setup}
con <- DBI::dbConnect(duckdb::duckdb())
dbWriteTable(con, "flights", nycflights13::flights)
flights_db <- tbl(con, "flights")
# Had to do it this way as for whatever reason using "copy_nycflights13" caused it to get stuck in a loop
```


1. What is distinct() translated to? How about head()?

```{r SQL translations}
flights_db |> 
  distinct(dest) |> 
  show_query()

# <SQL>
# SELECT DISTINCT dest
# FROM flights

flights_db |> 
  head(5) |> 
  show_query()

# <SQL>
# SELECT *
# FROM flights
# LIMIT 5
```

2. Explain what each of the following SQL queries do and try recreate them using dbplyr.

SELECT * 
FROM flights
WHERE dep_delay < arr_delay

SELECT *, distance / (air_time / 60) AS speed
FROM flights

```{r SQL to dplyr}
# The first selects all the columns from the flights dataset where departure delay is less than arrival delay. The dplyr equivalent would be:

flights_db |> 
  filter(dep_delay < arr_delay) |> 
  show_query()

# The second creates a new column called "speed" using the distance and air_time columns from the flights dataset. The dplyr equivalent would be:

flights_db |> 
  mutate(speed = distance / (air_time / 60)) |> 
  show_query()
```

