---
title: "R4DS_2E_iteration"
author: "Martin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Iteration - Modifying multiple columns

1. Practice your across() skills by:

  Computing the number of unique values in each column of palmerpenguins::penguins.

  Computing the mean of every column in mtcars.

  Grouping diamonds by cut, clarity, and color then counting the number of observations and computing the mean   of each numeric column.
  
```{r Number of unique values in each column}
palmerpenguins::penguins |>
  summarize(across(everything(), n_distinct))
```

```{r Mean of every column}
mtcars |> 
  summarize(across(everything(), mean))
# If any missing values were present we might have had to do the following.

mtcars |> 
  summarize(across(everything(), \(x) mean(x, na.rm = TRUE)))
```

```{r Diamonds breakdown}
diamonds |> 
  group_by(cut, clarity, color) |> 
  summarize(
    across(where(is.numeric), mean),
    n = n(),
    .groups = "drop"
    )
```

2. What happens if you use a list of functions in across(), but don’t name them? How is the output named?

```{r Unnamed listed functions in across}
df <- tibble(
  grp = sample(2, 10, replace = TRUE),
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

df |> 
  summarize(across(a:d, list(median, mean)))

df |> 
  summarize(across(a:d, list(median = median, mean = mean)))

# Each unnamed function takes the position its in for naming columns. So since `median` was the first unnamed function of the list, then output when naming columns treats it as `1` etc.
```

3. Adjust expand_dates() to automatically remove the date columns after they’ve been expanded. Do you need to embrace any arguments?

```{r Adjusting expand_dates}
# Lets create another argument with a default of `where(is.Date)` so that our function will now dynamically remove any columns that were originally in date format. Don't forget to embrace this new variable within our function
expand_dates <- function(df, date = where(is.Date)) {
  df |> 
    mutate(
      across({{ date }}, list(year = year, month = month, day = mday))
    ) |> 
    select(-{{ date }})
}

df_date <- tibble(
  name = c("Amy", "Bob"),
  this_is_a_date = ymd(c("2009-08-03", "2010-01-16"))
)

df_date |> 
  expand_dates()
```

4. Explain what each step of the pipeline in this function does. What special feature of where() are we taking advantage of?

```{r What does this function do?}
show_missing <- function(df, group_vars, summary_vars = everything()) {
  df |> 
    group_by(pick({{ group_vars }})) |> # First groups by the selected columns in `group_vars`
    summarize(
      across({{ summary_vars }}, \(x) sum(is.na(x))), # Calculates number of missing values for every column
      .groups = "drop" # Drops the `group_by` selection
    ) |>
    select(where(\(x) any(x > 0))) # Selects columns where any of the total missing flights is greater than 0
# Essentially its selecting all columns that can have a missing value due to cancelled or crashed flights etc.
# Anyway, the above showcases a special feature of `where` as it can use anonymous functions. Very handy.
}
nycflights13::flights |> show_missing(c(year, month, day))
```

