---
title: "Diamond sizes"
date: 2022-09-12
format: html
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(dplyr)

smaller <- diamonds |> 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds.
Only `r nrow(diamonds) - nrow(smaller)` are larger than 2.5 carats.
The distribution of the remainder is shown below:

```{r}
#| label: plot-smaller-diamonds
#| echo: false

smaller |> 
  ggplot(aes(x = carat)) + 
  geom_freqpoly(binwidth = 0.01)

lower_bound1 <- 0.55
upper_bound1 <- 0.65

lower_bound2 <- 0.66
upper_bound2 <- 0.75
```

The most striking features are that there is clearly some rounding going on when it comes to diamonds. Its hard to tell, but I reckon there is rounding to values like 1/3, 2/3, 1, etc.
For example there are only `r diamonds |> filter(between(carat, lower_bound1, upper_bound1)) |> count()` diamonds between `r lower_bound1` & `r upper_bound1` which is a difference of `r upper_bound1 - lower_bound1` compared to `r diamonds |> filter(between(carat, lower_bound2, upper_bound2)) |> count()` diamonds between `r lower_bound2` & `r upper_bound2` which is a difference of `r upper_bound2 - lower_bound2`